\pkg{cubble}: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2022-06-13 4-4-netcdf.Rmd
Add motivation example 5.5

The ERA5 dataset [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables. The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via the R package \pkg{ecmwfr} [@ecwmfr]. In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present. Here we specifially look at the information  contained on the variables *specific humidity* measuring the mass of water vapour per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level. Time-wise we restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once downloaded, the data can be read into a \code{cubble} as:

>>>>

The ERA5 dataset [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables. The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via the R package \pkg{ecmwfr} [@ecwmfr]. In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present. Here we specifically look at the information  contained on the variables *specific humidity* measuring the mass of water vapor per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level. The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2022 in the stratosphere in the Southern Hemisphere. Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once downloaded, the data can be read into a \code{cubble} as:

### 2022-06-13 arxiv/cubble-arxiv.tex
Add motivation example 5.5

The ERA5 dataset (\protect\hyperlink{ref-hersbach2020era5}{Hersbach et al. 2020}) provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables. The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} (\protect\hyperlink{ref-ecwmfr}{Hufkens, Stauffer, and Campitelli 2019}). In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present. Here we specifially look at the information contained on the variables \emph{specific humidity} measuring the mass of water vapour per kilogram of moist air and the \emph{geopotential} variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level. Time-wise we restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once downloaded, the data can be read into a \code{cubble} as:

>>>>

The ERA5 dataset (\protect\hyperlink{ref-hersbach2020era5}{Hersbach et al. 2020}) provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables. The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} (\protect\hyperlink{ref-ecwmfr}{Hufkens, Stauffer, and Campitelli 2019}). In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present. Here we specifially look at the information contained on the variables \emph{specific humidity} measuring the mass of water vapour per kilogram of moist air and the \emph{geopotential} variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level.The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2022 in the stratosphere. Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once downloaded, the data can be read into a \code{cubble} as:

### 2022-06-13 jss/cubble-jss.tex
Add motivation example 5.5

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}

@@

{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.007}

@@

{\centering \includegraphics[width=1\linewidth]{../figures/diagram-keynotes/diagram-keynotes.008}

@@

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.003}

@@

{\centering \includegraphics[width=1\linewidth]{../figures/illu-matching}

@@

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.004}

@@

{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{../figures/basic-manip}

@@

{\centering \includegraphics[width=1\linewidth]{../figures/basic-agg}

@@

{\centering \includegraphics[width=1\linewidth]{../figures/matching-map}

@@

{\centering \includegraphics[width=1\linewidth]{../figures/matching}

@@

The ERA5 dataset \citep{hersbach2020era5} is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF). The data can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} \citep{ecwmfr}. The \code{era5-pressure} data contains the \emph{specific humidity} and \emph{geopotential} variables on the 10 hPa pressure level on four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once downloaded, the data can be read into a cubble as:

@@

{\centering \includegraphics[width=1\linewidth]{../figures/netcdf}

@@

{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{../figures/linking-lower}

@@

{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{../figures/popup-mid}

@@

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.005}

>>>>

@@

{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}

@@

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/illu-matching}

@@

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}

@@

{\centering \includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/basic-manip}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/basic-agg}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/matching-map}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/matching}

@@

The ERA5 dataset \citep{hersbach2020era5} provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables. The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from \href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus Climate Data Store (CDS)} website or programmatically via the R package \pkg{ecmwfr} \citep{ecwmfr}. In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present. Here we specifically look at the information contained on the variables \emph{specific humidity} measuring the mass of water vapor per kilogram of moist air and the \emph{geopotential} variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level. The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2022 in the stratosphere in the Southern Hemisphere. Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once downloaded, the data can be read into a \code{cubble} as:

@@

{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/netcdf}

@@

{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/linking-lower}

@@

{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/popup-mid}

@@

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/PhD supervision/paper-cubble/figures/diagram-keynotes/diagram-keynotes.005}
